---
title: "Temperate Herbivorous Fish Feeding Ecology"
author: "Maya Lovelock-Dodos"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Desktop/Data_tidy_manuscript")
```

#Libraries
```{r, results='hide'}
library(tidyverse)
library(janitor)
library(mgcv)
library(cowplot)
library(scales)
library(geomtextpath)
library(ggtext)
library(ggplot2)
library(gridExtra)
library(ggthemes)
library(lme4)
```
```{r, publication theme, include=FALSE}
theme_Publication <- function(base_size=10, base_family="sans") {
  library(grid)
  library(ggthemes)
  (theme_foundation(base_size=base_size, base_family=base_family)
    + theme(plot.title = element_text(face = "bold",
                                      size = rel(1.2), hjust = 0.5, margin = margin(0,0,20,0)),
            text = element_text(),
            panel.background = element_rect(colour = NA),
            plot.background = element_rect(colour = NA),
            panel.border = element_rect(colour = NA),
            axis.title = element_text(face = "bold",size = rel(1.2)),
            axis.title.y = element_text(angle=90,vjust =2),
            axis.title.x = element_text(vjust = -0.2),
            axis.text = element_text(), 
            axis.line.x = element_line(colour="black"),
            axis.line.y = element_line(colour="black"),
            axis.ticks = element_line(),
            panel.grid.major = element_line(colour="#f0f0f0"),
            panel.grid.minor = element_blank(),
            legend.key = element_rect(colour = NA),
            legend.position = "bottom",
            legend.direction = "horizontal",
            legend.box = "vetical",
            legend.key.size= unit(0.5, "cm"),
            #legend.margin = unit(0, "cm"),
            legend.title = element_text(face="bold"),
            legend.text= element_text(face="italic", size = rel(1)),
            plot.margin=unit(c(10,5,5,5),"mm"),
            strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
            strip.text = element_text(face="bold")
    ))
  
}

scale_fill_Publication <- function(...){
  library(scales)
  discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#f87f01","#7fc97f","#ef3b2c","#feca01","#a6cee3","#fb9a99","#984ea3","#8C591D")), ...)
  
}

scale_colour_Publication <- function(...){
  library(scales)
  discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#f87f01","#7fc97f","#ef3b2c","#feca01","#a6cee3","#fb9a99","#984ea3","#8C591D")), ...)
  
}
```
# Q1: Do fish eat less at colder temperatures?

### Clean data
```{r Gastro-somatic Index}
maindat <- 
  read_csv("fishdata2.csv") |> 
  clean_names() |> 
  mutate(site = as.factor(site), 
         species = as.factor(species))
clean_dat <- 
  maindat |> 
  select(site, species, temp, weight_tot, fl, gsi, fult_k) 
```

### Data Exploration
Fork length and body weight with species and temperature groups pooled
```{r plot1, echo=FALSE}
clean_dat |> 
  ggplot(aes(log(fl), 
             log(weight_tot))) +
  geom_point()
```
```{r}
lw_mod <- lm(log(weight_tot)~log(fl)*species, data = clean_dat)
clean_dat_withcond <- 
clean_dat |> 
  mutate(est_w = exp(predict(lw_mod))) |> 
  mutate(resid_weight = weight_tot-est_w) 

```

seperate species fork length and weight
```{r plot2, echo=FALSE}
clean_dat_withcond|> 
  ggplot(aes(log(fl), 
             log(weight_tot), 
             col = species)) +
  geom_point() +
  geom_line(aes(y = log(est_w)))
```
```{r plot3, echo=FALSE}
clean_dat |> 
  ggplot(aes(x = I(0.01*fl^3), 
             y = weight_tot, 
             col = species)) +
  geom_point() +
  geom_smooth(method = "lm")
```
Temperature and gastosomatic index (or fullness) 
```{r plot4, echo=FALSE}
clean_dat |> 
  ggplot(aes(x = temp,
             y = gsi, 
             col = species)) +
  geom_point() +
  stat_smooth()
```
temperature and weight (g)
```{r plot5, echo=FALSE}
clean_dat |> 
  ggplot(aes(x = temp,
             y = weight_tot, 
             col = species)) +
  geom_point() +
  stat_smooth()
```

weight(g) and gsi by species
```{r}
clean_dat |> 
  ggplot(aes(x = weight_tot,
             y = gsi, 
             col = species)) +
  geom_point() +
  stat_smooth()
```

weight and gsi by temperature
```{r}
clean_dat |> 
  ggplot(aes(x = weight_tot,
             y = gsi, 
             col = temp)) +
  geom_point() +
  stat_smooth()
```


```{r}

lm1 <- lm(gsi ~ temp + species, data = clean_dat)
lm2 <- lm(gsi ~ temp + species*weight_tot, data = clean_dat) 
```
seems it's the weight driving the gsi, not the species ID, and _A. arctidens_ has the higher weight therefore higher gsi

```{r}
get_pvals <- function(mod){
  coefs <- data.frame(coef(summary(mod)))
  # use normal distribution to approximate p-value
  coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
  coefs
}

lmer(gsi ~ temp + species*weight_tot + (1|site), data = clean_dat) |> get_pvals()

jk<-lm(weight_tot~species, data = clean_dat)
summary(jk)
```

### GAMs
```{r GAMs}
mod1 <- gam(gsi~temp, data = clean_dat) 

mod2 <- gam(gsi~temp + site, data = clean_dat) 

mod3 <- gam(gsi~temp + weight_tot + site, data = clean_dat) 

mod4 <- gam(gsi~temp + species + weight_tot + site, data = clean_dat)

mod5 <- gam(gsi~temp + species*weight_tot + s(site, bs = 're'), data = clean_dat, method = 'REML')
mod6 <- gam(gsi~ s(temp, bs = "cr", k = 3) + s(weight_tot, bs = "cr") + species + s(site, bs = 're'), data = clean_dat, method = 'REML')

AIC(lm1, lm2, mod1, mod2, mod3, mod4, mod5, mod6) |> 
  arrange(desc(AIC))
#mod 6 wins
summary(mod6)
```

final plot for GSI (or gut fullness) using mod6.
```{r, echo=FALSE}
p1 <- clean_dat |> 
  mutate(fit12 = predict(mod6, newdata = data.frame(species = species, weight_tot = weight_tot, site=site, temp = 12))) |> 
  mutate(fit15 = predict(mod6, newdata = data.frame(species = species, weight_tot = weight_tot, site=site, temp = 15))) |> 
  mutate(fit17 = predict(mod6, newdata = data.frame(species = species, weight_tot = weight_tot, site=site, temp = 17))) |> 
  mutate(fit19 = predict(mod6, newdata = data.frame(species = species, weight_tot = weight_tot, site=site, temp = 19))) |> 
  ggplot(aes(x = weight_tot,
             y = gsi, 
             col = species)) +
  geom_point() +
  geom_textline(aes(y = fit12), label = "12°C", col = "#000033")+
  geom_textline(aes(y = fit15), label = "15°C", col = "#003399")+
  geom_textline(aes(y = fit17), label = "17°C", col = "#3399CC") +
  geom_textline(aes(y = fit19), label = "19°C", col = "#66CCFF") +
  theme(legend.title = element_text(size = 13)) +
  theme(legend.text = element_text(size = 10, face = "italic")) +
  labs(x = "Total weight (g)", 
       y = "GSI") +
  ggtitle("Trend of gastrosomatic index (GSI) with weight and temperature (°C)")+ 
  theme(plot.title = element_text(size = 5))

p1 + scale_colour_Publication()+ theme_Publication()
```
# Q2 What are they eating? Gut content analysis (GCA)
### note: Multivariate analysis done using PRIMER

```{r}
GCAdat <- 
  read_csv("fdata.all.csv") |> 
  clean_names() |> 
  mutate(site = as.factor(site), 
         species = as.factor(species),
         lipids = as.factor(lipids),
         fult_k = as.numeric(fult_k),
         month = as.factor(month))
summary(GCAdat)
```
```{r}
gca <- 
  GCAdat |> 
  select(id,month,temp, species,fl, weight_tot,lipids, fult_k, phylum,spp,perc_composition) 


summary(gca)
a2 <- gca[c(1:330, 419:902, 1123:1408, 1497:1606),]
summary(a2)
c2 <- gca[c(331:418, 903:1122, 1409:1496, 1607:1650),]
summary(c2)
```

Final plot used to visualise diet composition of species at different temperatures

```{r, echo=FALSE}
GCAplot<- gca %>% 
  filter(phylum != 'Arthropoda') %>%
  filter(phylum != 'Cnidaria') %>%
  filter(phylum != 'Bryozoa') %>%
  ggplot(aes(fill=phylum, y=perc_composition, x=temp)) + 
  geom_bar(position="fill", stat="identity")+
  scale_fill_manual(values = c("Chlorophyta" = "darkolivegreen3",
                                "Ochrophyta" = "lightgoldenrod4",
                                "Porifera" = "darkorange",
                                "Rhodophyta" = "palevioletred"
                                )) +
  theme(plot.title=element_text(size=15)) +
  facet_grid(vars(species)) +
  labs(x= "Temperature (°C)",
       y = "Composition") +
  ggtitle ("Composition of macroalgae")
GCAplot + scale_colour_Publication()+ theme_Publication()
```
# Q3 Is the condition of the sampled fish changing with temperature/seasons

### 3.1: investigating the residuals of length and weight to get an indication of condition
```{r, echo=FALSE}
clean_dat_withcond |> 
  ggplot(aes(x = temp, 
             y = resid_weight,
             col = species)) +
  geom_point() +
  stat_smooth()
```

```{r, echo=FALSE}
clean_dat_withcond |> 
  ggplot(aes(x = weight_tot, 
             y = resid_weight, 
             col = species)) +
  geom_point() +
  stat_smooth()
```

```{r}
lm(resid_weight ~ temp , data = clean_dat_withcond)


lm(resid_weight ~ temp + species*weight_tot, data = clean_dat_withcond) |> summary()
mod7 <- gam(resid_weight ~ s(temp, bs = "cr", k = 5) + s(weight_tot, bs = "cr") + 
              s(site, bs = 're'), data = clean_dat_withcond, method = 'REML')

mod8 <- gam(resid_weight ~ s(temp, bs = "cr", k = 5) + s(weight_tot, bs="re"),
              data = clean_dat_withcond, method = 'REML') 
```

```{r}
mod7
clean_dat_withcond |> 
  mutate(fit12 = predict(mod7, newdata = data.frame(species = species, weight_tot = weight_tot, site=site, temp = 12))) |> 
  mutate(fit15 = predict(mod7, newdata = data.frame(species = species, weight_tot = weight_tot, site=site, temp = 15))) |> 
  mutate(fit17 = predict(mod7, newdata = data.frame(species = species, weight_tot = weight_tot, site=site, temp = 17))) |> 
  mutate(fit19 = predict(mod7, newdata = data.frame(species = species, weight_tot = weight_tot, site=site, temp = 19))) |> 
  ggplot(aes(x = weight_tot,
             y = resid_weight, 
             col = species)) +
  geom_point() +
  geom_line(aes(y = fit12), col = "black")+
  geom_line(aes(y = fit15), col = "blue")+
  geom_line(aes(y = fit17), col = "red") +
  geom_line(aes(y = fit19), col = "orange") +
  theme_cowplot() +
  labs(x = "Total weight", 
       y = "Resid weight (g)")
```

### Fulton's K conditional factor as an indicator of body condition
```{r, echo=FALSE}
clean_dat_withcond |> 
  ggplot(aes(x = temp, 
             y = fult_k, 
             col = species)) +
  geom_point() +
  stat_smooth() +
  labs(x="Temperature °C",
       y="Fulton's K")+
  ggtitle("Fulton's K with Temperature")
```

```{r, echo=FALSE}
clean_dat_withcond |> 
  ggplot(aes(x = weight_tot, 
             y = fult_k, 
             col = species)) +
  geom_point() +
  stat_smooth()
```

### GLMs
```{r}
lm(fult_k ~ species, data = clean_dat_withcond) |> summary()

lm(fult_k ~ species + temp, data = clean_dat_withcond) |> summary()

lm(fult_k ~ species + temp*weight_tot, data = clean_dat_withcond) |> summary()

lm(fult_k ~ species + temp*weight_tot + site, data = clean_dat_withcond) |> summary()

lm8 <- lm(fult_k ~ species*weight_tot +temp + site, data = clean_dat_withcond)
lm9 <- lm(fult_k ~ species + weight_tot + site, data = clean_dat_withcond)
lm10 <- lm(fult_k ~ species + weight_tot + site + temp, data = clean_dat_withcond)

summary(lm9)
summary(lm10)

lm(fult_k ~ species*site, data = clean_dat_withcond) |> summary()

AIC(lm8, lm9, lm10) |> 
  arrange(desc(AIC))
```
```{r, echo=FALSE}
p <- clean_dat_withcond |> 
  mutate(fit = predict(lm9)) |> 
  ggplot(aes(x = weight_tot, 
             y = fult_k, 
             col = species, 
             lty = site)) +
  geom_point() +
  geom_line(aes(y = fit)) +
  labs(x = "Total weight (g)",
       y = "Fulton's K") +
  ggtitle("Changes in Fulton's K across site and species") +
  theme(legend.title = element_text(size = 10)) +
  theme(legend.text = element_text(size = 10))
p + scale_colour_Publication()+ theme_Publication()
```

